On the presentation of orthogonal groups over polynomial rings

### 2019-05-02 dl-paper.tex

Denote by $P_\alpha(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_\alpha(0)$ and the elements $x_{-\alpha}(mX)$, $m \in M$. The following lemma shows that this subgroup is sufficiently large.
\begin{lemma} \label{Pstar-large} The subgroup $P_\alpha(*)$ contains the image of $K(A[X], M[X])$ in $\St(\Phi, A[X, X^{-1}])$. \end{lemma}
\begin{proof} Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$. Invoking~\eqref{Z5} of~\cref{Zrels} with $s = Xf$, $\eta = 1$ we conclude from this that $z_{\pm\alpha}(Xf, \xi)$ also belong to $P_{\alpha}(*)$. Thus, we have shown that $P_\alpha(*)$ contains $\St(\Phi, A[X], M[X])$.

Now choose $\gamma$ to be arbitrary root orthogonal to $\alpha$. By~\cref{Kgen-strong} we are left to prove that $c_{\pm \gamma}(f, X\xi)$ lies in $P_\alpha(*)$. 
But $c_{\pm \gamma}(f, X\xi) = x_{\pm \gamma}(f) \cdot z_{\pm \gamma}(-f, -X\xi)$ is a product of two generators of type 3. \end{proof}

@@

\begin{lemma}\label{P0_normal} The subgroup $P_\alpha(0)$ is normal in $P_\alpha(*)$. In particular, there is an exact sequence split by the map $m \mapsto x_{-\alpha}(mX)$:
\[\xymatrix{1 \ar[r] & P_\alpha(0) \ar[r] & P_\alpha(*) \ar[r] & M^+ \ar@<5pt>@{-->}[l] \ar[r] & 1}.\] \end{lemma}
\begin{proof} We need to verify that the conjugate by $x_{-\alpha}(mX)$ to every generator of $P_\alpha(0)$ listed in~Definition~\ref{defP0} still belongs to $P_\alpha(0)$.
The assertion is obvious for the generators of type (3) and (4).

@@

Finally, in the case $\alpha + \beta \in \Phi$ (only possible for the generator of type (1)) we have:
\[ z_{\beta}(Xf, \xi) ^ {x_{-\alpha}(mX)} = x_{-\alpha} (mX^2f\xi ) \cdot x_{-\alpha-\beta} (N_{\beta,\alpha}\cdot mX^2f\xi^2) \cdot z_{\beta}(Xf, \xi), \]
the latter expression, clearly, is a product of generators of type $4, 1, 1$.
\end{proof}

@@

Let us verify the assertion for the generator of type (4).
Choose $\beta\in \Phi$ such that $\alpha+\beta \in \Phi$ and invoke~\eqref{Z5} of~\cref{Zrels} substituting $s = Xf$, $\eta = X$, $\xi = aX^{-1}$:
\begin{multline} \nonumber z_{-\alpha}(X^2f, aX^{-1}) = x_{-\alpha-\beta}(\epsilon Xf) \cdot x_{-\beta}(-af) \cdot x_{\beta}(X^2 f a) \cdot x_{-\alpha}(X^2f) \cdot \\
 \cdot z_{-\alpha-\beta}(-\epsilon Xf, -\epsilon a) \cdot x_{\alpha+\beta}(-\epsilon Xf a^2) \cdot x_{\alpha}(- f a^2) \cdot z_{-\beta}(f a, -X), \end{multline}
where $\epsilon = N_{-\alpha-\beta,\beta}$. 
Clearly, the expression in the right-hand side is a product of generators of type 1, 2, 1, 1, 1, 1, 2, 2.

It remains to verify the assertion for the remaining generator $z_\alpha(f, gX)$ of type 2.
Substituting in~\eqref{Z5} of~\cref{Crels} $s := -\epsilon f$, $\xi = -\epsilon g$, $\eta=X$ we obtain that
\begin{multline} \nonumber z_{\alpha+\beta}(-\epsilon fX, -\epsilon \xi) = x_\alpha( -f) x_{-\beta}(- f \xi) \cdot x_{\beta}( X^2f \xi) \cdot x_{\alpha+\beta}(-X\epsilon f) \cdot \\
 \cdot z_\alpha(f,  X\xi) \cdot x_{-\alpha}(X^2f\xi^2) \cdot x_{-\alpha-\beta}(\epsilon Xf \xi^2) \cdot z_{-\beta}(f\xi, -X), \end{multline}
where $\epsilon = N_{\alpha,\beta}$. 
We can express $z_{\alpha}(f, X\xi)$ through other terms and obtain that
\begin{multline} \nonumber z_\alpha(f, X\xi) = x_{\alpha+\beta}(\epsilon Xf) \cdot x_{\beta}(-X^2 f \xi) \cdot x_{-\beta}(f \xi) \cdot x_\alpha(f) \cdot \\ 
 \cdot z_{\alpha+\beta}(-\epsilon X f, -\epsilon \xi) \cdot z_{-\beta}(-f\xi, -X) \cdot x_{-\alpha-\beta}(-\epsilon X f \xi^2) \cdot x_{-\alpha}(-X^2 f\xi^2) \cdot \end{multline}
The factors in the right-hand side are generators of type 1, 1, 2, 2, 1, 2, 1, 4. 
It is clear, that conjugation by $x_{\alpha}(aX^{-1})$ acts trivially on the fourth factor $x_\alpha(f)$, 
while for the other factors we have already verified that their conjugates belong to $P_{\alpha}(0)$. \end{proof}

@@

\begin{corollary} For $g \in P_\alpha(m)$ holds $S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am) \in P(\alpha, \frac{m}{1 + am})$.
Consequently one has \[ S_\alpha(a+b, g) = S_\alpha\left(b,\ S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)\right)\cdot h_\alpha(1+am). \]\end{corollary}
\begin{proof}
Suppose $g \in P_\alpha(m)$, then $g = g_0 \cdot x_{-\alpha}(mX)$ for some $g_0 \in P_\alpha(0)$ for which we already know that $S_\alpha(a, g_0) \in P(\alpha, 0)$.
We need to verify that $h = S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)$ lies in $P\left(\alpha, \frac{m}{1 + am}\right)$, but from Lemmas~\ref{P0_conj}--\ref{lem:Tulenbayev-formula} and the antisymmetry of Steinberg symbols we obtain that:
\begin{multline} h = {}^{x_\alpha(aX^{-1})}\!g_0 \cdot \left(x_{\alpha}(aX^{-1}) \cdot x_{-\alpha}(mX)\right) \cdot x_{\alpha}\left(-\frac{aX^{-1}}{1+am}\right) \cdot \{X, 1+am\} \cdot h_\alpha^{-1}(1 + am) = \\
= S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \in P_\alpha\left(\frac{m}{1+am}\right). \end{multline}
The second assertion can be verified in a straightforward manner. \end{proof}

\begin{lemma} If $\alpha - \beta \in \Phi$ then for all $g \in \St(\Phi, R[X], XM[X])$ holds $S_\alpha(a, g) \in P(\beta, 0)$. \end{lemma}
\begin{proof}
It suffices to verify the assertion for the generators $z_\gamma(Xf, g)$, $\gamma \in \Phi$.
By our assumption $(\alpha, \beta) \geq 0$. 
\end{proof}

\subsection{The proof of the injectivity theorem}

 \begin{comment}
\begin{lemma} if $\alpha \perp \beta$ then $g \in P(\alpha, m) \cap P(\beta, m')$ implies $S_\alpha(a, g) \in P(\beta, m')$. \end{lemma}
\begin{proof}
 Lol
\end{proof}
\end{comment}

>>>>

Denote by $P_\alpha(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_\alpha(0)$ and the elements $x_{-\alpha}(mX)$, $m \in M$. The following lemma shows that this subgroup is sufficiently large.
\begin{lemma} \label{Pstar-large} The subgroup $P_\alpha(*)$ contains the image of $K(A[X], M[X])$ in $\St(\Phi, A[X, X^{-1}])$. \end{lemma}
\begin{proof} Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$. 
Invoking~\eqref{Z5} of~\cref{Zrels} with $s = Xf$, $\eta = 1$ we obtain that $z_{\pm\alpha}(Xf, \xi)$ also belong to $P_{\alpha}(*)$. Thus, we have shown that $P_\alpha(*)$ contains $\St(\Phi, A[X], M[X])$.

Now choose $\gamma$ to be arbitrary root orthogonal to $\alpha$. By~\cref{Kgen-strong} we are left to prove that $c_{\pm \gamma}(f, X\xi)$ lies in $P_\alpha(*)$. 
But $c_{\pm \gamma}(f, X\xi) = x_{\pm \gamma}(f) \cdot z_{\pm \gamma}(-f, -X\xi)$ is a product of two generators of $P_\alpha(0)$ of type 3. \end{proof}

@@

\begin{lemma}\label{P0_normal} The subgroup $P_\alpha(0)$ is normal in $P_\alpha(*)$. In particular, there is an exact sequence, which is split by the map $m \mapsto x_{-\alpha}(mX)$:
\[\xymatrix{1 \ar[r] & P_\alpha(0) \ar[r] & P_\alpha(*) \ar[r] & M^+ \ar@<5pt>@{-->}[l] \ar[r] & 1}.\]
\end{lemma}
\begin{proof} We need to verify that the conjugate by $x_{-\alpha}(mX)$ to the generators of $P_\alpha(0)$ listed in~Definition~\ref{defP0} belongs to $P_\alpha(0)$.
The assertion is obvious for the generators of type (3) and (4).

@@

Finally, in the case $\alpha + \beta \in \Phi$ (only possible for the generator of type (1)) we have:
\[ z_{\beta}(Xf, \xi) ^ {x_{-\alpha}(mX)} = x_{-\alpha} (mX^2f\xi ) \cdot x_{-\alpha-\beta} (N_{\beta,\alpha}\cdot mX^2f\xi^2) \cdot z_{\beta}(Xf, \xi). \]
It is clear that the latter expression is a product of generators of type $4, 1, 1$.
\end{proof}

@@

Let us verify the assertion for the generator of type (4).
Choose $\beta\in \Phi$ such that $\alpha+\beta \in \Phi$. 
Substituting $s = Xf$, $\eta = X$, $\xi = aX^{-1}$ and $\alpha = -\alpha - \beta$ into~\eqref{Z5} of~\cref{Zrels} we obtain that
\begin{multline} \nonumber z_{-\alpha}(X^2f, aX^{-1}) = x_{-\alpha-\beta}(\epsilon Xf) \cdot x_{-\beta}(-af) \cdot x_{\beta}(X^2 f a) \cdot x_{-\alpha}(X^2f) \cdot \\
 \cdot z_{-\alpha-\beta}(-\epsilon Xf, -\epsilon a) \cdot x_{\alpha+\beta}(-\epsilon Xf a^2) \cdot x_{\alpha}(- f a^2) \cdot z_{-\beta}(f a, -X), \end{multline}
where $\epsilon = N_{-\alpha-\beta,\beta}$. 
Clearly, the expression in the right-hand side is a product of generators of type 1, 2, 1, 1, 1, 1, 2, 2.

It remains to verify the assertion for the remaining generator $z_\alpha(f, X\xi)$ of type 2.
Substituting in~\eqref{Z5} of~\cref{Crels} $s = -\epsilon f$, $\xi = -\epsilon \xi$, $\eta=X$ and expressing $z_\alpha(f, X\xi)$ through other terms we obtain that
\begin{multline} \nonumber z_\alpha(f, X\xi) = x_{\alpha+\beta}(\epsilon Xf) \cdot x_{\beta}(-X^2 f \xi) \cdot x_{-\beta}(f \xi) \cdot x_\alpha(f) \cdot \\ 
 \cdot z_{\alpha+\beta}(-\epsilon X f, -\epsilon \xi) \cdot z_{-\beta}(-f\xi, -X) \cdot x_{-\alpha-\beta}(-\epsilon X f \xi^2) \cdot x_{-\alpha}(-X^2 f\xi^2), \end{multline}
where $\epsilon = N_{\alpha,\beta}$.  
The factors in the right-hand side are generators of type 1, 1, 2, 2, 1, 2, 1, 4. 
It is clear, that conjugation by $x_{\alpha}(aX^{-1})$ acts trivially on the fourth factor $x_\alpha(f)$, 
while for the other factors we have already verified that their conjugates belong to $P_{\alpha}(0)$. \end{proof}

@@

\begin{corollary} For $g \in P_\alpha(m)$ holds $S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am) \in P(\alpha, \frac{m}{1 + am})$.
Consequently one has \[ S_\alpha(a+b, g) = S_\alpha\left(b,\ S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)\right)\cdot h_\alpha(1+am). \]\end{corollary}
\begin{proof}
Suppose $g \in P_\alpha(m)$, then $g = g_0 \cdot x_{-\alpha}(mX)$ for some $g_0 \in P_\alpha(0)$ for which we already know that $S_\alpha(a, g_0) \in P(\alpha, 0)$.
We need to verify that $h = S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am)$ lies in $P\left(\alpha, \frac{m}{1 + am}\right)$, but from Lemmas~\ref{P0_conj}--\ref{lem:Tulenbayev-formula} and the antisymmetry of Steinberg symbols we obtain that:
\begin{multline} \nonumber h = {}^{x_\alpha(aX^{-1})}\!g_0 \cdot \left(x_{\alpha}(aX^{-1}) \cdot x_{-\alpha}(mX)\right) \cdot x_{\alpha}\left(-\frac{aX^{-1}}{1+am}\right) \cdot \{X, 1+am\} \cdot h_\alpha^{-1}(1 + am) = \\
= S_\alpha(a, g_0) \cdot x_{-\alpha}\left(\frac{mX}{1+am}\right) \cdot \langle -a, m\rangle_\alpha \in P_\alpha\left(\frac{m}{1+am}\right). \end{multline}
The second assertion can be verified in a straightforward manner. \end{proof}

\begin{lemma} If $\alpha - \beta \in \Phi$ then for all $g \in \St(\Phi, R[X], XM[X])$ holds $S_\alpha(a, g) \in P_\beta(0)$. \end{lemma}
\begin{proof}
It suffices to verify the assertion for the generators $z_\gamma(Xf, g)$, $\gamma \in \Phi$.
By our assumption $(\alpha, \beta) \geq 0$. 
\end{proof}

\subsection{The proof of the injectivity theorem}
We denote by $B$ the ring $M[X] + A[X^{-1}]$ considered as a graded subring of $A[X, X^{-1}]$.
Denote by $\overline{G}^{\geq 0}_M$ the image of the map $\St(\Phi, R[X], M[X]) \to \St(\Phi, A[X, X^{-1}])$ and by
 $G^{\leq 0}$ the group $\St(\Phi, A[X^{-1}])$.

\begin{lemma} if $\alpha \perp \beta$ then $g \in P(\alpha, m) \cap P(\beta, m')$ implies $S_\alpha(a, g) \in P(\beta, m')$. \end{lemma}
\begin{proof}
 Lol
\end{proof}
